Particle Interception as a Function of Reynolds Number and Collector Density

### 2020-08-27 code/analysis.Rmd
finished sed trap and velocity data

In the interest of documenting our analytic methods, I've decided to make one comprehensive, hopefully lay-readable document including all steps.

@@

pumpfiles <- list.files("data/peristaltic pumps/")
trapfiles <- list.files("data/sediment traps/")

@@

pumpdata <- lapply(pumpfiles, function(x) read_csv(paste0("data/peristaltic pumps/",x)))

@@

So now we have a nice long table of our pump data (time x concentration) stratified by run (i.e., treatment), height, and upstream/downstream location.

@@

pump %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_jitter(alpha = .5, height = 0) +
  facet_wrap(~date)

@@

pumpfinal <- pump %>%

@@

length(unique(pumpfinal$date))

@@

metadata <- read_csv("data/run_metadata.csv")

sum(unique(pumpfinal$date) %in% metadata$date)

@@

pumpfinal <- left_join(pumpfinal, metadata, by = "date")

@@

pumpfinal %>%
  filter(growth_days==0) %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_jitter(alpha = .5, height = 0) +
  facet_wrap(~date)

@@

pumpfinal %>%
  filter(growth_days==0, date != 190417, dowel_density != "0232") %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_jitter(alpha = .5, height = 0) +
  facet_grid(pump_freq~dowel_density)

@@

>>>>

In the interest of documenting our analytic methods, I've decided to make one comprehensive, hopefully lay-readable document including all steps from our raw laboratory data to the model results. I won't be able to include the turbulence analyses in time for our meeting (8/27), but that required Laurel's Matlab script anyways.

@@

pumpfiles <- list.files("../data/peristaltic pumps/")
trapfiles <- list.files("../data/sediment traps/")

@@

pumpdata <- lapply(pumpfiles, function(x) read_csv(paste0("../data/peristaltic pumps/",x)))

@@

So now we have a long table of our pump data (time $\times$ concentration) stratified by run (i.e., treatment), height, and upstream/downstream location.

@@

pump %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_point(alpha = .5, size = .75) +
  facet_wrap(~date)

@@

pump <- pump %>%

@@

length(unique(pump$date))

@@

metadata <- read_csv("../data/run_metadata.csv")

sum(unique(pump$date) %in% metadata$date)

@@

pump <- left_join(pump, metadata, by = "date")

@@

pump %>%
  filter(growth_days==0) %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_point(alpha = .5, size = .75) +
  geom_rect(data = filter(pump, date == 190417), 
                          fill = NA, colour = "red", xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf) +
  facet_wrap(~date) +
  scale_x_continuous(breaks = c(0,3000,6000), limits = c(0,6000))

@@

pumpfinal <- pump %>%
  filter(growth_days==0, date != 190417, dowel_density != "0232")

pumpfinal %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_point(alpha = .5, size = .75) +
  facet_grid(pump_freq~dowel_density) +
  scale_x_continuous(breaks = c(0,3000,6000), limits = c(0,6000))

@@

There are certainly no trends jumping out immediately, which isn't necessarily worrisome since exponential decay is hard to compare by eye. More on that later though.

### Sediment trap data

Data import and wrangling follows a similar structure to the pump data.

trapdata <- lapply(trapfiles, function(x) read_csv(paste0("../data/sediment traps/",x)))

names(trapdata) <- trapdate

x <- trapdata

tidytrap <- lapply(seq_along(x), function(i) {
  select(x[[i]], station = 1, pre = contains("pre"), post = contains("post"), sed = contains("sed")) %>%
  mutate(date = names(x)[i]) %>%
  mutate_at(vars(pre,post,sed,date), as.numeric)
  }
  )

trap <- bind_rows(tidytrap) %>%
  filter(!is.na(sed))

trap
```

So, across 21 runs, we had 186 samples. We were set up to collect 9 per run, but traps broke, filters slipped and spilled sediment, et cetera. $186/21 =$ `r 186/21`, so our success rate was actually pretty good, although some of those measurements were probably erroneous despite being measured. Let's see what we're working with and then we can maybe try to remove outliers.

lm(sed~station, data = trap) %>% summary()

So, downstream traps tended to have less settling. This challenges our assumption of uniformity throughout the test section and might justify the use of a spline, as Justin and Candace had suggested.

trap %>%
  group_by(date) %>%
  mutate(mean = mean(sed)) %>%
  arrange(mean) %>%
  ggplot(aes(x = factor(date, levels = unique(date)), y = sed)) +
  geom_boxplot() +
  labs(x = "date", y = "average sediment mass (g)") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

On the whole, the sediment data seems pretty normal and independent. Let's zoom in to just the 12 runs used in the main analysis.

left_join(trap, metadata, by = "date") %>%
  filter(date %in% pumpfinal$date) %>%
  group_by(date) %>%
  mutate(mean = mean(sed)) %>%
  arrange(mean) %>%
  ggplot(aes(x = factor(date, levels = unique(date)), y = sed)) +
  geom_boxplot() +
  labs(x = "date", y = "average sediment mass (g)") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

A pretty broad range, without any clear outliers to knock out, at least by my sense of it. The whiskers are limited to $1.5 \cdot \text{IQR}$ each (both up and down), so while there are 4 points outside that range for their respective runs, they're definitely all within the range of the dataset as a whole.

### Velocity

We estimated Reynolds number from a flow velocity experiment in the open channel conducted in early 2019.

![Flow Velocity Regression](../pics/frequency_velocity_boxplots.jpeg)

The data and code for this analysis is MIA at the moment, but should turn up soon.
